---
title: "Chapter-9"
output: 
  ioslides_presentation: 
    keep_md: yes
    widescreen: yes
---

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Load packages and functions

```{r}
# (vegan must be loaded after ade4 to avoid some conflicts)
library(ade4)
library(vegan)
library(gclus)
library(ape)
```

## Load additional functions
```{r}
# (files must be in the working directory)
source("evplot.R")
source("cleanplot.pca.R")
source("PCA.R")
source("CA.R")
```

## Import data
```{r}
# (files must be in the working directory)
spe <- read.csv("DoubsSpe.csv", row.names=1)
env <- read.csv("DoubsEnv.csv", row.names=1)
spa <- read.csv("DoubsSpa.csv", row.names=1)
# Remove empty site 8
spe <- spe[-8,]
env <- env[-8,]
spa <- spa[-8,]
```

## CA of the raw species dataset (original species abundances)
```{r}
# Compute CA
spe.ca <- cca(spe)
spe.ca
summary(spe.ca)		# default scaling 2
summary(spe.ca, scaling=1)
```

## Plot eigenvalues and % of variance for each axis
```{r}
(ev2 <- spe.ca$CA$eig)
```
## plot 
```{r}
evplot(ev2)
```


## CA biplots
```{r}
par(mfrow=c(1,2))
# Scaling 1: sites are centroids of species
plot(spe.ca, scaling=1, main="CA fish abundances - biplot scaling 1")
# Scaling 2 (default): species are centroids of sites
plot(spe.ca, main="CA fish abundances - biplot scaling 2")
```
## A posteriori projection of environmental variables in a CA

```{r}
# The last plot produced (CA scaling 2) must be active
plot(spe.ca, main="CA fish abundances - biplot scaling 2")
(spe.ca.env <- envfit(spe.ca, env))
plot(spe.ca.env)
# Plot significant variables with a different colour
plot(spe.ca.env, p.max=0.05, col=3)
```

## Species data table ordered after the CA result

```{r}
vegemite(spe, spe.ca)
```
## CA using CA() function

```{r}
spe.CA.PL <- CA(spe)
biplot(spe.CA.PL, cex=1)

# Ordering of the data table following the first CA axis
# The table is transposed, as in vegemite() output
summary(spe.CA.PL)
t(spe[order(spe.CA.PL$F[,1]), order(spe.CA.PL$V[,1])])
```

## NMDS applied to the fish species - Bray-Curtis distance matrix

```{r}
spe.nmds <- metaMDS(spe, distance="bray")
spe.nmds
spe.nmds$stress
plot(spe.nmds, type="t", main=paste("NMDS/Bray - Stress =", 
	round(spe.nmds$stress,3)))
```






## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```

