---
title: "Chapter-9"
author: "Joey and Sean"
output: 
  ioslides_presentation: 
    keep_md: yes
    smaller: yes
    widescreen: yes
---

## Plan for today
- PCA
- PCoA
- NMDS
- Correspondence analysis

## Load packages and functions

```{r warning=FALSE, message=FALSE}
# (vegan must be loaded after ade4 to avoid some conflicts)
library(ade4)
library(vegan)
library(gclus)
library(ape)
```

## Load additional functions
```{r}
# (files must be in the working directory)
source("evplot.R")
source("cleanplot.pca.R")
source("PCA.R")
source("CA.R")
```

## NMDS
- the objective is to plot dissimilar objects far apart in the ordination space and similar objects close to one another.
- can use any distance matrix
- can deal with missing data


## NMDS applied to the varespec data: vegetation lichen pastures

```{r}
data(varespec) ## 44 species, lots of zeros, so  not amenable to PCA
str(varespec)
```

## metaMDS - a wrapper function

- Wraps several recommended procedures into one command:
 + takes raw data and performs 'Wisconsin double standardization' (so abundance isn't influencing similarity)
 + calculates specified dissimilarity matrix
 + runs vegan function `monoMDS` many times with random starts, stopping when it finds two similar configurations with minimum stress
 + rotates solution so largest variation of site score is on first axis
 + other details in [vegan tutor](http://cc.oulu.fi/~jarioksa/opetus/metodi/vegantutor.pdf)
 
## dissimilarity indices

- `metaMDS` automatically standardizes and then calculates specified dissimilarity index
- `vegdist` will take a matrix of sites (rows) and variables/species (columns) and calculate specied dissimilarity index, outputs class `dist`
```{r}
varespec.kul <- vegdist(varespec, method="kulczynski")
str(varespec.kul)
```

## Running NMDS

```{r}
varespec.nmds.bray <- metaMDS(varespec, distance="bray", trace=FALSE, trymax=100) ## trace = FALSE, won't show all the outputs, trymax= number of starts
varespec.nmds.bray
```


# ```{r}
# spe.nmds <- metaMDS(spe, distance="bray")
# spe.nmds
# spe.nmds$stress
# plot(spe.nmds, type="t", main=paste("NMDS/Bray - Stress =", 
# 	round(spe.nmds$stress,3)))
# ```



## plotting NMDS

```{r}
plot(varespec.nmds.bray, type="t")
```


## With many variables/sites, ordination plots can quickly become overwhelming
- some built-in function to help include:
 + `make.cepnames` (shortens latin names to 6 char variables)
 + `orditorp` (reduces overlapping points and variable names)
 + `ordilabel` 

## evaluating NMDS mapping

```{r}
stressplot(varespec.nmds.bray) ## plots the observed disimilarity values vs. their ordination distance. If NMDS is a good representation of actual values, you'll see a good fit
```


## goodness of fit
```{r}
gof <- goodness(varespec.nmds.bray) # gof for each site. 
plot(varespec.nmds.bray, type="t", main="goodness of fit") # larger circles represent plots that don't have a strong fit with original disimilarity matrix.
points(varespec.nmds.bray, display="sites", cex=gof*100)
```

## Comparing ordinations

- Comparing different ordinations can be difficult because of slightly
different orientation and scaling. Procrustes rotation using `procrustes` allows comparison
```{r}
varespec.nmds.eu <- metaMDS(varespec, distance="eu", trace=FALSE, trymax=100) # use euclidean distance - probably not a good choice for most community analyses
pro <- procrustes(varespec.nmds.bray, varespec.nmds.eu)
```

## Comparing ordinations

```{r}
plot(pro, cex=1.5)
```

## Comparing ordinations

```{r}
plot(pro, kind=2) # shows the shift in sites between two ordinations.
```





## Import data
```{r}
# (files must be in the working directory)
spe <- read.csv("DoubsSpe.csv", row.names=1)
env <- read.csv("DoubsEnv.csv", row.names=1)
spa <- read.csv("DoubsSpa.csv", row.names=1)
# Remove empty site 8
spe <- spe[-8,]
env <- env[-8,]
spa <- spa[-8,]
```

## Intro to CA
- CA is well suited to the analysis of species abundance data without pre-transformation. 

- data submitted to CA must be frequencies or frequency-like, dimensionally homogeneous and non-negative (as is the case of species counts or presence–absence data)

-  raw data are transformed into a matrix Q, of cell-by-cell contributions to the pearson χ 2 statistic

- get an ordination, where the χ 2 distance is preserved among sites instead of the Euclidean distance. The χ 2 distance is not influenced by double zeros

## How CA differs from PCA
- In both PCA and CA the weights are derived by eigenanalysis
- In PCA, the matrix of species abundances is transformed into a matrix of covariances or correlations, each abundance value being replaced by a measure of its correlation (or covariance) with other abundances in other quadrats
- In CA, the abundance data is transformed to a chi-square statistic, which is used to depict the degree associations among sites depart from independence
-  the chi-square metric in CA preserves ecological distance by modeling differences in associations rather than abundances of single species. 

- the system of weights used to score sites or quadrats is derived from a metric of species associations, and the more these associations depart from independence, the further separated final scores will be


## CA of the raw species dataset (original species abundances)
```{r}
# Compute CA
spe.ca <- cca(spe)
spe.ca
summary(spe.ca)		# default scaling 2
summary(spe.ca, scaling=1) 
## things to note: the first axis has a very large eigenvalue. In CA, a value over 0.6 indicate a very strong gradient in the data. Note that the eigenvalues are the same in both scalings. The scaling affects the eigenvectors, not the eigenvalues. 
```
## Scaling options
- when scaling = 1 the distances among objects approximate their x2 distances (i.e. object points close together are similar in their species frequencies). Any object near the point representing a species is likely to contain a high contribution of that species. 

- when scaling = 2, ordination of species. Species points close to one another are likely to have similar relative frequencies among objects. 



## Plot eigenvalues and % of variance for each axis
```{r}
ev2 <- spe.ca$CA$eig
```
## plot 
```{r}
evplot(ev2)
```
-things to note: the first axis is extremely dominant. 

## CA biplots
```{r}
par(mfrow=c(1,2))
# Scaling 1: sites are centroids of species
plot(spe.ca, scaling=1, main="CA fish abundances - biplot scaling 1")
# Scaling 2 (default): species are centroids of sites
plot(spe.ca, main="CA fish abundances - biplot scaling 2")
```
## A posteriori projection of environmental variables in a CA
-envfit finds vectors or factor averages of environmental variables. [...] The projections of points onto vectors have maximum correlation with corresponding environmental variables, and the factors show the averages of factor levels”
	
## CA biplot (scaling 2) of the Doubs fish abundance data with a posteriori projection of environmental variables

```{r}
# The last plot produced (CA scaling 2) must be active
plot(spe.ca, main="CA fish abundances - biplot scaling 2")
(spe.ca.env <- envfit(spe.ca, env))
plot(spe.ca.env)
# Plot significant variables with a different colour
plot(spe.ca.env, p.max=0.05, col=3)
```

## Species data table ordered after the CA result

```{r}
vegemite(spe, spe.ca)
```
## CA using CA() function

```{r}
spe.CA.PL <- CA(spe)
biplot(spe.CA.PL, cex=1)

# Ordering of the data table following the first CA axis
# The table is transposed, as in vegemite() output
summary(spe.CA.PL)
t(spe[order(spe.CA.PL$F[,1]), order(spe.CA.PL$V[,1])])
```
## useful links
- [PCA vs CA?](http://stats.stackexchange.com/questions/70533/using-principal-components-analysis-vs-correspondence-analysis)
